Analytical Predictions of Exoplanet Obliquities Generated by Planet-Disk Interactions

### 2020-02-24 writeups/Su_disk_disp.tex
move most plots into writeup folder, almost finished Dong first pass

    \documentclass[
        fleqn,
        usenatbib,
        referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
        caption,
        subcaption,
    }
    \usepackage[T1]{fontenc}
    \captionsetup{compatibility=false}

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\uv}[1]{\hat{\bm{#1}}}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

Since $\p{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\ref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\ref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2},\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{The solid lines show the fractional areas of each of the zones
    $A_{\rm i}(\eta) / 4\pi$ as given by Eqs.~\eqref{se:area_ward}. The colored
    dashed lines correspond to small $\eta$ approximations used in
    Appendix~\ref{s:ad_approx}. The colored dashed lines for $\eta > \eta_{\rm
    c}$ are the effective values of $A_{\rm II}, A_{\rm III}$ for $\eta >
    \eta_{\rm c}$, denoting the points that would flow into either area under
    adiabatic decrease of $\eta$ from $\eta > \eta_{\rm c}$. The vertical black
    dashed lines correspond to $\eta = \eta_{\rm c}$ [Eq.\eqref{eq:etac}] and
    the values of $\eta$ for which $A_{\rm II}$ is maximized ($\eta_{\rm \max,
    II}$) and for which $A_{\rm III}$ is minimized ($\eta_{\rm \min, III}$,
    Eq.~\eqref{eq:eta_minIII}). The horizontal black dashed line illustrates the
    fraction of phase space occupied by zone II at separatrix appearance.
    }\label{fig:eq_areas}
\end{figure}

@@

Beginning at the last separatrix crossing, the final enclosed phase space area
$A_{\rm f}$ will be conserved for all time. As $\eta \to 0$, trajectories circulate at
constant obliquity $\theta_{\rm f}$, related to $A_{\rm f}$ by
\begin{equation}
    2\pi\p{1 - \cos \theta_{\rm f}} = A_{\rm f}. \label{eq:qfaf}
\end{equation}

$A$ is not conserved when the trajectory encounters the separatrix. Howover, its
change is easily understood \citep{henrard1982}. In essence, when the trajectory
crosses the separatrix, it continues to evolve adjacent to the separatrix. So if
a separatrix crossing results in a zone I trajectory, the resultant $A$ can be
approximated by integrating Eq.~\eqref{eq:a_oint} along the upper leg of the
separatrix. Pictorally, this can be seen in the bottom plots of
Fig.~\ref{fig:ad_21}.

\subsubsection{Governing Principles: Probabilistic Separatrix Crossing}

When a trajectory experiences separatrix crossing, it transitions into nearby
zones probabilistically. This process is well understood by
\citealp{henrard1982,henrard1987}. Their results may be simply summarized as
follows: if a zone $i$ is shrinking while adjacent zones $j, k$ are expanding
such that the sum of their areas is constant, the probability of transition from
zone $i$ to zone $j$ is given
\begin{align}
    \Pr\p{i \to j} = -\frac{\pd{A_{\rm j}}{t}}{ \pd{A_{\rm i}}{t}}
        = -\frac{\pd{A_{\rm j}}{\eta}}{ \pd{A_{\rm i}}{\eta}},\\
    \Pr\p{i \to k}
        = -\frac{\pd{A_{\rm k}}{\eta}}{ \pd{A_{\rm i}}{\eta}}.\label{eq:henrard_hop}
\end{align}
Note that $\Pr \p{i \to j} + \Pr\p{i \to k} = 1$. This can be used directly in
conjunction with Eqs.~\eqref{se:area_ward} to understand for what initial
conditions each track can be observed and with what probabilities.

As a particular example, consider a system in zone II in panel (d) of
Fig.~\ref{fig:eq_1contours}. As $\eta decreases, zone II$ will shrink while
zones I and III will expand until the trajectory. Suppose the trajectory
exits zone II at some $\eta_\star$, then the probability of a II $\to$ I
transition is $\Pr\p{\rm II \to I} = -\frac{\dot{A}_{\rm I}}{\dot{A}_{\rm II}}$,
while a II $\to$ III transition occurs with probability $\Pr\p{\rm II \to III} =
-\frac{\dot{A}_{\rm III}}{\dot{A}_{\rm II}}$.

@@

Returning to the evolution of $\uv{s}$, we can classify trajectories by the
separatrix encounters experienced. Initially, in the $\eta > \eta_{\rm c}$ regime,
only zones II and III exist. Conversely, at the end of the simulation when $\eta
\to 0$, only zones I and III exist. Most simply, one might expect four sequences
of transitions between zones (we call these dynamical ``tracks'') to manifest:
from one of $\z{I, III}$ to one of $\z{II, III}$, where III $\to$ III is a
trivial case (no separatrix encounter). In addition to these four tracks, a
fifth track III $\to$ II $\to$ I is observed. Below, we will describe the
evolution of $A$ throughout each of the five tracks, as well as determine the
initial conditions and probabilities associated with each track:
\begin{enumerate}
    \item II $\to$ I --- An example trajectory following this track is depicted
        in Fig.~\ref{fig:ad_21}. $\uv{s}$ starts librating about CS2 in zone II,
        enclosing some initial phase space area $A_{\rm i}$. The trajectory is
        able to librate without separatrix encounter until $A_{\rm
        II}(\eta_\star) = A_{\rm i}$. As the trajectory transitions to a
        circulating trajectory in zone I immediately bordering the separatrix,
        it will encompass $-A_{\rm I}(\eta_\star)$ phase space area. The final
        $\theta_{\rm f}$ is then given by Eq.~\eqref{eq:qfaf}. An analytical
        approximation to $\theta_{\rm f}$ is derived in
        Appendix~\ref{s:ad_approx} and is
        \begin{equation}
            \cos \theta_{\rm f, II \Rightarrow I} \approx
                \p{\frac{\pi \theta_{\rm sp, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{\rm sp, i}^2}{4}.\label{eq:qf_21_approx}
        \end{equation}

        This transition can only occur when the initial condition begins in zone
        II, requiring $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$. Note $A_{\rm II}(\eta_{\rm c})$ is
        given in closed form in \citet{ward2004I} but appears to be incorrect.
        Then, at some known $\eta_\star$ satisfying $A_{\rm II}(\eta_\star) =
        A_{\rm i}$, the trajectory is ejected from the separatrix following
        Eq.~\eqref{eq:henrard_hop}. Note that since $\pd{A_{\rm i}}{\eta} < 0$
        everywhere, while $\pd{A_{\rm II}}{\eta} > 0$ at all possible
        $\eta_\star$ for an initial condition starting in zone II, this track
        always has nonzero probability.

    \item II $\to$ III --- An example trajectory following this track is
        depicted in Fig.~\ref{fig:ad_23}. The only difference from the previous
        track is that, upon separatrix encounter, the trajectory follows the
        circulating trajectory in zone III bordering the separatrix, upon
        which it will encompass $A_{\rm f} = A_{\rm I}(\eta_\star) + A_{\rm
        II}(\eta_\star)$. The final obliquity is still given by
        Eq.~\eqref{eq:qfaf}, and the analytical approximation derived in
        Appendix~\ref{s:ad_approx} is
        \begin{equation}
            \cos \theta_{\rm f, II \Rightarrow III} \approx
                \p{\frac{\pi \theta_{\rm sp, i}^2}{16}}^2 \cot I
                    - \frac{\theta_{\rm sp, i}^2}{4}.\label{eq:qf_23_approx}
        \end{equation}

        Again, this track can only occur when $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$, but a
        further constraint arises when we consider the transition probability
        given by Eq.~\eqref{eq:henrard_hop}. Upon examination of
        Fig.~\ref{fig:eq_areas}, it is clear that $\pd{A_{\rm III}}{\eta} > 0$ for
        many $\eta$. Call
        \begin{equation}
            \eta_{\min, III} \equiv \argmin A_{\rm III}(\eta)
                \label{eq:eta_minIII},
        \end{equation}
        which is labeled in Fig.~\ref{fig:eq_areas}, then if $\eta_\star >
        \eta_{\rm \min, III}$ then $\Pr_{\rm II \to III} < 0$. This is
        understood as a forbidden transition, and so II $\to$ III is only a
        permitted dynamical track if $\eta_\star <, \eta_{\rm \min, III}$.

    \item III $\to$ I --- This track is depicted in Fig.~\ref{fig:ad_31}.
        The trajectory encounters the separatrix when $A_{\rm I}(\eta_\star) +
        A_{\rm II}(\eta_\star) = A_{\rm i}$, upon which it transitions to a zone
        I trajectory enclosing $A_{\rm f} = -A_{\rm I}$. Then, as always, the
        final obliquity is given by Eq.~\eqref{eq:qfaf}.

        This track can only occur if $A_{\rm i} > A_{\rm II}(\eta_{\rm c})$, but is also
        constrained by requiring its $A_{\rm i}$ is sufficiently small it will
        encounter the separatrix (if it is too large, it will never encounter
        the separatrix, which is a III $\to$ III transition). This is written
        $A_{\rm i} < \max A_{\rm I} + A_{\rm II} = 4\pi - \min A_{\rm III}$.

        Note that since $\pd{A_{\rm I}}{\eta} < 0$ always while $\pd{A_{\rm
        III}}{\eta} > 0$ for all accessible $\eta_{\star}$, this track is always
        permitted.

    \item III $\to$ II $\to$ I --- This track is depicted in
        Fig.~\ref{fig:ad_321}. The first separatrix encounter occurs at $\eta_1$
        when $A_{\rm I}(\eta_1) + A_{\rm II}(\eta_1) = A_{\rm i}$, upon which
        the trajectory moves into zone II enclosing intermediate phase space
        area $A_{\rm m} = A_{\rm II}(\eta_1)$. Examining
        Fig.~\ref{fig:eq_areas}, there exist values $A$ for which $A_{\rm
        II}(\eta) = A$ has multiple solutions, and $\eta_1 > \eta_{\rm II,
        \max}$ as labeled. These are the only values for which the III $\to$ II
        $\to$ I transition is permitted. Thus, as $\eta$ continues to decrease,
        a second $\eta_2$ value exists for which $A_{\rm m} = A_{\rm
        II}(\eta_2)$, upon which the trajectory is ejected to zone I and $A_f =
        -A_{\rm I}(\eta_2)$. The final obliquity again then obeys
        Eq.~\eqref{eq:qfaf}.

@@

    \item III $\to$ III --- This track depicts the trivial case where no
        separatrix encounter ever occurs, and $A$ is constant throughout the
        evolution. This constraint equates to $A_{\rm i} > \max A_{\rm I} + A_{\rm
        II}$, whereupon no separatrix encounter occurs at all, and $A_{\rm f} =
        A_{\rm i}$,
        which for $\eta_{\rm i} \to \infty, \eta_{\rm f} \to 0$ results in $\theta =
        \theta_{\rm sd, i}$.

        Note that a small deviation from $\theta = \theta_{\rm sd, i}$ occurs
        because we use finite $\eta_{\rm i}$, as hinted at in Section~\ref{sss:a_evo}.
        This is because for $\theta_{\rm sd, i} \gtrsim \pi/2$, trajectories are
        better described as librating about CS3. Nevertheless, we can still
        approximate $A_{\rm i}$ to good accuracy as the solid angle enclosed by
        libration about CS3 with polar angle $2\pi + \theta_3 - \theta_2 -
        \theta_{\rm sd, i}$. This recovers the correct transfer function
        $\theta_{\rm f}\p{\theta_{\rm sd, i}}$ for the III $\to$ III regime, as
        can be seen visibly in Fig.~\ref{fig:3_ensemble_20_35}.
\end{enumerate}
In summary, the five regimes of track possibilities in increasing order of
$A_{\rm i}$
are:
\begin{enumerate}
    \item $A_{\rm i} \in \s{0, A_{\rm II}\p{\eta_{\rm \min, III}}}$ --- II $\to$
        III and II $\to$ I both possible.

    \item $A_{\rm i} \in \s{A_{\rm II}\p{\eta_{\rm \min, III}}, A_{\rm
        II}(\eta_{\rm c})}$ --- II $\to$ I only.

    \item $A_{\rm i} \in \s{A_{\rm II}(\eta_{\rm c}), A_{\rm II, \max}}$ ---
        III $\to$ I and III $\to$ II $\to$ I both possible.

    \item $A_{\rm i} \in \s{A_{\rm II, \max}, \max A_{\rm I} + A_{\rm II}}$ ---
        III $\to$ I only.

>>>>

%TODO
% rework 3->2->1 ??? part, probably also the 3 -> 3 part
% Fig 7 caption lol
    \documentclass[
        fleqn,
        usenatbib,
        referee,
    ]{mnras}
    \usepackage{
        amsmath,
        amssymb,
        newtxtext,
        newtxmath,
        graphicx,
        ae, aecompl,
        booktabs,
        caption,
        subcaption,
    }
    \usepackage[T1]{fontenc}
    \captionsetup{compatibility=false}

    \newcommand*{\rd}[2]{\frac{\mathrm{d}#1}{\mathrm{d}#2}}
    \newcommand*{\rtd}[2]{\frac{\mathrm{d}^2#1}{\mathrm{d}#2^2}}
    \newcommand*{\pd}[2]{\frac{\partial#1}{\partial#2}}
    \newcommand*{\pdil}[2]{\partial#1 / \partial#2}
    \newcommand*{\md}[2]{\frac{\mathrm{D}#1}{\mathrm{D}#2}}
    \newcommand*{\at}[1]{\left.#1\right|}
    \newcommand*{\abs}[1]{\left|#1\right|}
    \newcommand*{\ev}[1]{\langle#1\rangle}
    \newcommand*{\bm}[1]{\boldsymbol{\mathbf{#1}}}
    \newcommand*{\uv}[1]{\hat{\bm{#1}}}
    \newcommand*{\p}[1]{\left(#1\right)}
    \newcommand*{\s}[1]{\left[#1\right]}
    \newcommand*{\z}[1]{\left\{#1\right\}}
    \DeclareMathOperator*{\argmin}{argmin}
    \DeclareMathOperator*{\argmax}{argmax}
    \DeclareMathOperator*{\med}{med}

@@

Since $\p{\phi, \cos \theta}$ form a set of canonical variables, the integral
$\oint \cos \theta\;\mathrm{d}\phi$ along a trajectory is an adiabatic invariant
(see Section~\ref{s:ad}). The unsigned areas $\p{\abs{\int \cos \theta
\;\mathrm{d}\phi}}$ of the three zones (as defined in
Fig.~\ref{fig:eq_1contours}) can be computed analytically
\citep{henrard1987,ward2004I}. Define
\begin{subequations}
    \begin{align}
        z_0 &= \eta\cos I, &
        \chi &= \sqrt{-\frac{\tan^3\theta_4}{\tan I} - 1},\\
        \rho &= \chi \frac{\sin^2 \theta_4\cos \theta_4}{
            \chi^2 \cos^2\theta_4 + 1},&
        T &= 2\chi \frac{\cos \theta_4}{
            \chi^2 \cos^2\theta_4 - 1}.
    \end{align}
\end{subequations}
The areas for $\eta < \eta_{\rm c}$ are given by
\begin{subequations}\label{se:area_ward}
    \begin{align}
        A_{\rm I} &= 2\pi\p{1 - z_0} - \frac{A_2}{2}, \label{eq:A1}\\
        A_{\rm II} &= 8\rho + 4\arctan T - 8z_0 \arctan \frac{1}{\chi},
            \label{eq:A2}\\
        A_{\rm III} &= 2\pi\p{1 + z_0} - \frac{A_2}{2}\label{eq:A3}.
    \end{align}
\end{subequations}
These are plotted as a function of $\eta$ in Fig.~\ref{fig:eq_areas}. Note that
the zones are not formally defined for $\eta > \eta_{\rm c}$ since the
separatrix disappears, but a natural extension exists: evolve an initial phase
space point $p$ under adiabatic decrease of $\eta$ until the separatrix appears
at $\eta = \eta_{\rm c}$, then identify $p$ with the zone it is in at $\eta_{\rm
c}$. Since phase space area is conserved under adiabatic evolution, this
extension implies $A_{\rm i}\p{\eta > \eta_{\rm c}} = A_{\rm i}(\eta_{\rm c})$.
The boundary between these extended zones is denoted by the dashed black line in
the top left panel of Fig.~\ref{fig:eq_1contours}, where otherwise no separatrix
exists.
\begin{figure}
    \centering
    \includegraphics[width=0.5\textwidth]{../initial/99_misc/1_areas.png}
    \caption{The solid lines show the fractional areas of each of the zones
    $A_{\rm i}(\eta) / 4\pi$ as given by Eqs.~\eqref{se:area_ward}. The colored
    dashed lines correspond to small $\eta$ approximations used in
    Appendix~\ref{s:ad_approx}. The colored dashed lines for $\eta > \eta_{\rm
    c}$ are the effective values of $A_{\rm II}, A_{\rm III}$ for $\eta >
    \eta_{\rm c}$, denoting the points that would flow into either area under
    adiabatic decrease of $\eta$ from $\eta > \eta_{\rm c}$. The vertical black
    dashed lines correspond to $\eta = \eta_{\rm c}$ [Eq.\eqref{eq:etac}] and
    the values of $\eta$ for which $A_{\rm II}$ is maximized ($\eta_{\rm \max,
    II}$) and for which $A_{\rm III}$ is minimized ($\eta_{\rm \min, III}$,
    Eq.~\eqref{eq:eta_minIII}). The horizontal black dashed line illustrates the
    fraction of phase space occupied by zone II at separatrix appearance.
    }\label{fig:eq_areas}
\end{figure}

@@

Beginning at the last separatrix crossing, the final enclosed phase space area
$A_{\rm f}$ will be conserved for all time. As $\eta \to 0$, trajectories
circulate about $\hat{l}$ at constant obliquity $\theta_{\rm f}$, related to
$A_{\rm f}$ by
\begin{equation}
    2\pi\p{1 - \cos \theta_{\rm f}} = A_{\rm f}. \label{eq:qfaf}
\end{equation}

The enclosed phase space area is not conserved when the trajectory encounters
the separatrix. Howover, the change is easily understood \citep{henrard1982}. In
essence, when the trajectory crosses the separatrix, it continues to evolve
adjacent to the separatrix. So if a separatrix crossing results in a zone I
trajectory (see Fig.~\ref{fig:eq_1contours}), the new area can be approximated
by integrating Eq.~\eqref{eq:a_oint} along the upper leg of the separatrix.
Pictorally, this can be seen in the bottom panels of Fig.~\ref{fig:ad_21}.

\subsubsection{Governing Principle: Probabilistic Separatrix Crossing}

When a trajectory experiences separatrix crossing, it transitions into nearby
zones probabilistically. This process is studied by \citet{henrard1982} and
\citet{henrard1987}. Their results may be summarized as follows: if zone $i$
is shrinking while adjacent zones $j, k$ are expanding such that the sum of
their areas is constant, the probabilities of transition from zone $i$ to zones
$j$ and $k$ are given by
\begin{subequations}\label{eq:henrard_hop}
    \begin{align}
        \Pr\p{i \to j} = -\frac{\pdil{A_{\rm j}}{\eta}}{\pdil{A_{\rm i}}{\eta}},
                \\
        \Pr\p{i \to k} = -\frac{\pdil{A_{\rm k}}{\eta}}{\pdil{A_{\rm i}}{\eta}}.
    \end{align}
\end{subequations}
Note that $\Pr \p{i \to j} + \Pr\p{i \to k} = 1$.
Eqs.~\eqref{eq:henrard_hop} can be used directly in conjunction with
Eqs.~\eqref{se:area_ward} to understand for what initial conditions each track
can be observed and with what probabilities.

As an example, consider a system in zone II in panel (d) of
Fig.~\ref{fig:eq_1contours}. As $\eta decreases, zone II$ will shrink while
zones I and III will expand until the trajectory crosses the separatrix. Suppose
the trajectory exits zone II at some $\eta_\star$, then the probability of the
II $\to$ I transition is $\Pr\p{\rm II \to I} = -\dot{A}_{\rm I} / \dot{A}_{\rm
II}$, while the II $\to$ III transition occurs with probability $\Pr\p{\rm II
\to III} = -\dot{A}_{\rm III} / \dot{A}_{\rm II}$.

@@

Returning to the evolution of $\uv{s}$, we can classify trajectories by the
sequence of separatrix encounters. Initially, in the $\eta > \eta_{\rm c}$
regime, only zones II and III exist; as $\eta \to 0$, only zones I and III exist
(see Fig.~\ref{fig:eq_1contours}). There are five distinct evolutionary tracks:
\begin{enumerate}
    \item II $\to$ I (see Fig.~\ref{fig:ad_21} for an example). The spin axis
        $\uv{s}$ initially circulates in zone II (see a), and then starts
        librating about CS2 as $\eta$ decreases (b), enclosing some initial
        phase space area $A_{\rm i}$. This libration continues until the
        separatrix expands (due to decreasing $\eta$) to ``touch'' the
        trajectory (see c), at which $A_{\rm II}(\eta_\star) = A_{\rm i}$.
        As $\hat{s}$ to a
        circulating trajectory in zone I immediately bordering the separatrix,
        it will encompass $-A_{\rm I}(\eta_\star)$ phase space area. The final
        obliquity $\theta_{\rm f}$ is then given by Eq.~\eqref{eq:qfaf}, with
        $A_{\rm f} = -A_{\rm I}\p{\eta_\star}$. An analytical approximation to
        $\theta_{\rm f}$ is derived in Appendix~\ref{s:ad_approx} and is
        \begin{equation}
            \p{\cos \theta_{\rm f}}_{\rm II \to I} \approx
                \p{\frac{\pi \theta_{\rm sd, i}^2}{16}}^2 \cot I
                    + \frac{\theta_{\rm sd, i}^2}{4}.\label{eq:qf_21_approx}
        \end{equation}
        The transition probability is
        \begin{equation}
            \Pr\p{\mathrm{II} \to \mathrm{I}} = -\p{
                \frac{\pdil{A_{\rm I}}{\eta}}{\pdil{A_{\rm II}}{\eta}}}
                    _{\eta = \eta_{\star}}.
        \end{equation}

        This track can only occur when the initial condition begins in zone II,
        requiring $A_{\rm i} < A_{\rm II}(\eta_{\rm c})$, where
        \begin{equation}
            A_{\rm II}\p{\eta_c} = 4\pi\s{1 - \p{1 + \tan^{2/3}I}^{-3/2}},
        \end{equation}
        is given by Eq.~\eqref{eq:A2} evaluated at $\eta = \eta_c$
        \citep{ward2004I}.

        Also, note that since $\pdil{A_{\rm I}}{\eta} < 0$ everywhere, while
        $\pdil{A_{\rm II}}{\eta} > 0$ at all possible $\eta_\star$ for an
        initial condition starting in zone II, this track always has nonzero
        probability.

    \item II $\to$ III (see Fig.~\ref{fig:ad_23}) This track is similar to the
        II $\to$ I track; the only difference is that, upon separatrix
        encounter, the trajectory follows the circulating trajectory in zone III
        bordering the separatrix, upon which it will encompass area $A_{\rm
        I}(\eta_\star) + A_{\rm II}(\eta_\star) = A_{\rm f}$. The final
        obliquity is still given by Eq.~\eqref{eq:qfaf}, and the analytical
        approximation derived in Appendix~\ref{s:ad_approx} is
        \begin{equation}
            \p{\cos \theta_{\rm f}}_{\rm II \to III} \approx
                \p{\frac{\pi \theta_{\rm sd, i}^2}{16}}^2 \cot I
                    - \frac{\theta_{\rm sd, i}^2}{4}.\label{eq:qf_23_approx}
        \end{equation}
        The transition probability is
        \begin{equation}
            \Pr\p{\mathrm{II} \to \mathrm{III}} = -\p{
                \frac{\pdil{A_{\rm III}}{\eta}}{\pdil{A_{\rm II}}{\eta}}}
                    _{\eta = \eta_{\star}}.
        \end{equation}

        Again, this track can only occur when $A_{\rm i} < A_{\rm II}(\eta_{\rm
        c})$, but a further constraint arises when we consider the transition
        probability. Upon examination of Fig.~\ref{fig:eq_areas}, it is clear
        that $\pd{A_{\rm III}}{\eta} > 0$ for a large range of $\eta$, which
        would give a negative transition probability---implying a forbidden
        transition. Define
        \begin{equation}
            \eta_{\min, III} \equiv \argmin A_{\rm III}(\eta)
                \label{eq:eta_minIII},
        \end{equation}
        which is labeled in Fig.~\ref{fig:eq_areas}. Thus, the II $\to$ III
        track is permitted only if $\eta_\star <, \eta_{\rm \min, III}$.

    \item III $\to$ I (see Fig.~\ref{fig:ad_31}) The trajectory encounters the
        separatrix when $A_{\rm I}(\eta_\star) + A_{\rm II}(\eta_\star) = A_{\rm
        i}$, upon which it transitions to a zone I trajectory enclosing $A_{\rm
        f} = -A_{\rm I}$. The final obliquity is again given by
        Eq.~\eqref{eq:qfaf}.

        This track can only occur if $A_{\rm i} > A_{\rm II}(\eta_{\rm c})$, but
        is also constrained by requiring $A_{\rm i}$ be sufficiently small so
        that it will encounter the separatrix (if $A_{\rm i}$ is too large, it
        will never encounter the separatrix, and we simply have a III $\to$ III
        transition). This condition is $A_{\rm i} < \max \p{A_{\rm I} + A_{\rm
        II}} = 4\pi - \min \p{A_{\rm III}}$.

        Note that since $\pdil{A_{\rm I}}{\eta} < 0$ while $\pdil{A_{\rm
        III}}{\eta} > 0$ for all accessible $\eta_{\star}$, this track is always
        permitted.

    \item III $\to$ II $\to$ I (see Fig.~\ref{fig:ad_321}) The first separatrix
        encounter occurs at $\eta_1$ when $A_{\rm I}(\eta_1) + A_{\rm
        II}(\eta_1) = A_{\rm i}$, upon which the trajectory moves into zone II
        enclosing intermediate phase space area $A_{\rm m} = A_{\rm
        II}(\eta_1)$. Examining Fig.~\ref{fig:eq_areas}, there exist values $A$
        for which $A_{\rm II}(\eta) = A$ has multiple solutions, and $\eta_1 >
        \eta_{\rm II, \max}$ as labeled. These are the only values for which the
        III $\to$ II $\to$ I transition is permitted. Thus, as $\eta$ continues
        to decrease, a second $\eta_2$ value exists for which $A_{\rm m} =
        A_{\rm II}(\eta_2)$, upon which the trajectory is ejected to zone I and
        $A_f = -A_{\rm I}(\eta_2)$. The final obliquity again then obeys
        Eq.~\eqref{eq:qfaf}.

@@

    \item III $\to$ III --- This track depicts the trivial case where no
        separatrix encounter occurs, and $A$ is constant throughout the
        evolution ($A_{\rm f} = A_{\rm i}$). This requires $A_{\rm i} > \max
        \p{A_{\rm I} + A_{\rm II}}$. In the limit of $\eta_{\rm i} \to \infty$
        and $ \eta_{\rm f} \to 0$ we have $\theta_{\rm f} = \theta_{\rm sd, i}$.

        Note that a small deviation from $\theta = \theta_{\rm sd, i}$ occurs
        because we use finite $\eta_{\rm i}$, as hinted at in
        Section~\ref{sss:a_evo}. This is because for $\theta_{\rm sd, i} \gtrsim
        \pi/2$, trajectories are better described as librating about CS3.
        Nevertheless, we can still approximate $A_{\rm i}$ to good accuracy as
        the solid angle enclosed by libration about CS3 with polar angle $2\pi +
        \theta_3 - \theta_2 - \theta_{\rm sd, i}$. This recovers the correct
        transfer function $\theta_{\rm f}\p{\theta_{\rm sd, i}}$ for the III
        $\to$ III regime, as can be seen visibly in
        Fig.~\ref{fig:3_ensemble_20_35}.
\end{enumerate}
In summary, starting from an initial condition with phase space area $A_{\rm i}$
at $\eta = \eta_{\rm i} \gg 1$, the five evolutionary tracks are:
\begin{enumerate}
    \item $A_{\rm i} \in \s{0, A_{\rm II}\p{\eta_{\rm \min, III}}}$. Both the II
        $\to$ III and the II $\to$ I tracks are possible.

    \item $A_{\rm i} \in \s{A_{\rm II}\p{\eta_{\rm \min, III}}, A_{\rm
        II}(\eta_{\rm c})}$. Only the II $\to$ I track.

    \item $A_{\rm i} \in \s{A_{\rm II}(\eta_{\rm c}), A_{\rm II, \max}}$.
        Both the III $\to$ I and III $\to$ II $\to$ I  are possible.

    \item $A_{\rm i} \in \s{A_{\rm II, \max}, \max \p{A_{\rm I} + A_{\rm II}}}$.
        Only the III $\to$ I track.

    \item $A_{\rm i} > \max \p{A_{\rm I} + A_{\rm II}}$. Only the III $\to$ III
        track.
\end{enumerate}
In all cases, the corresponding ranges for $\theta_{\rm sd,i}$ can be computed via
Eq.~\eqref{eq:ai_qsd}. The boundaries between these ranges are overplotted in
Fig.~\ref{fig:ad_ensemble}, where they can be seen to agree well with the
numerical results.
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo21.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo21_subplots.png}
    \end{subfigure}
    \caption{Fiducial simulation following the $A_{\rm II} \to A_{\rm I}$
    transition. An initial $\theta_{\rm sd, i} = 0.3\;\mathrm{rad} \approx
    17.2^\circ$ was used, as well as $\epsilon = 3 \times 10^{-4}$. Top plot
    upper panel: plot of $\cos \theta(t)$ (green) in an example simulation.
    Overlaid are the locations of Cassini State 2 (red), and upper and lower
    bounds on the separatrix (black). Note that the trajectory successfully
    tracks CS2 to an obliquity $\theta \approx 90^\circ$. Light vertical blue
    lines denote times portrayed in bottom plot. Top plot lower panel: plot of
    the enclosed separatrix area obtained by integrating the simulated
    trajectory (blue dots) and adiabatic theory (red line). Even though the
    enclosed area jumps at separatrix crossing, the resultant $A$ is still
    accurately predicted. Lower plot: snapshots in the $\p{\cos \theta, \phi}$
    space of the simulation trajectories for the vertical blue lines in the top
    plots. These times correspond to the start of the simulation, the appearance
    of the separatrix, two panels depicting the separatrix crossing process, and
    a final snapshot at $\eta = 10^{-3.5}$. The green line denotes a full
    circulation or libration cycle at the selected time, including an arrow for
    directionality. The separatrix at the start/end of the portrayed cycle are
    shown in solid/dashed black lines respectively. Also labeled is CS2 at the
    start of the cycle (red). Finally, the enclosed phase space area is shaded
    in grey ($A > 0$) and red ($A < 0$).}\label{fig:ad_21}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo23.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo23_subplots.png}
    \end{subfigure}
    \caption{Same as Fig.~\ref{fig:ad_21} but for the $A_2 \to A_3$ track.
    $\theta_{\rm sd, i} = 0.3\;\mathrm{rad}$ and $\epsilon = 3.01 \times
    10^{-4}$.}\label{fig:ad_23}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo31.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo31_subplots.png}
    \end{subfigure}
    \caption{Same as Fig.~\ref{fig:ad_21} but for the $A_{\rm III} \to A_{\rm I}$
    track. $\theta_{\rm sd, i} = 89.1^\circ$, and $\epsilon = 3 \times
    10^{-4}$.}\label{fig:ad_31}
\end{figure}
\begin{figure}
    \centering
    \begin{subfigure}{\columnwidth}
        \centering
        \includegraphics[width=0.5\textwidth]{../initial/2_toy2/3testo321.png}
    \end{subfigure}
    \begin{subfigure}{\columnwidth}
        \centering
